### 2022-12-11 sentiment_eda.Rmd

---
title: "Final Paper Part 1"
author: "Sam Kalman and Charles Hutchinson"
date: "2022-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = F}
# Load Package
library(syuzhet)
library(tidyverse)
```

```{r, message = F}
# Load and Join Bay Area Fire Tweets

# 4680 Tweets by 7 Variables
bay_area_tweets <- readRDS("bay_area_tweets.RDS")
# 3807 Users by 8 Variables
bay_area_users <- readRDS("bay_area_users.RDS")
colnames(bay_area_users)[1] <- "author_id"

bay_area_bert <- read_csv("bay_area_sentiment.csv")

bay_area_tweets <- cbind(bay_area_tweets, 
                         bay_area_bert$label, 
                         bay_area_bert$score)

colnames(bay_area_tweets)[8] <- "bert_sent"
colnames(bay_area_tweets)[9] <- "bert_score"

# 3294 Unique Users
bay_area_users <- bay_area_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
bay <- inner_join(bay_area_tweets, bay_area_users, by = "author_id") %>%
  mutate(fire = "Bay Area Fire")
```

```{r, message = F}
# Load and Join Camp Fire Tweets

# 50469 Tweets by 7 Variables
camp_tweets <- readRDS("camp_tweets.RDS")
# 44376 Users by 8 Variables
camp_users <- readRDS("camp_users.RDS")
colnames(camp_users)[1] <- "author_id"

camp_bert <- read_csv("camp_sentiment.csv")

camp_tweets <- cbind(camp_tweets, 
                         camp_bert$label, 
                         camp_bert$score)

colnames(camp_tweets)[8] <- "bert_sent"
colnames(camp_tweets)[9] <- "bert_score"

# 26353 Unique Users
camp_users <- camp_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
camp <- inner_join(camp_tweets, camp_users, by = "author_id") %>%
  mutate(fire = "Camp Fire")
```

```{r, message = F}
# Load and Join Dixie Fire Tweets

# 26139 Tweets by 7 Variables
dixie_tweets <- readRDS("dixie_tweets.RDS")
# 21230 Users by 8 Variables
dixie_users <- readRDS("dixie_users.RDS")
colnames(dixie_users)[1] <- "author_id"

dixie_bert <- read_csv("dixie_sentiment.csv")

dixie_tweets <- cbind(dixie_tweets, 
                         dixie_bert$label, 
                         dixie_bert$score)

colnames(dixie_tweets)[8] <- "bert_sent"
colnames(dixie_tweets)[9] <- "bert_score"

# 11438 Unique Users
dixie_users <- dixie_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
dixie <- inner_join(dixie_tweets, dixie_users, by = "author_id") %>%
  mutate(fire = "Dixie Fire")
```

```{r, message = F}
# Load and Join Tubbs Fire Tweets

# 3530 Tweets by 7 Variables
tubbs_tweets <- readRDS("tubbs_tweets.RDS")
# 2826 Users by 8 Variables
tubbs_users <- readRDS("tubbs_users.RDS")
colnames(tubbs_users)[1] <- "author_id"

tubbs_bert <- read_csv("tubbs_sentiment.csv")

tubbs_tweets <- cbind(tubbs_tweets, 
                         tubbs_bert$label, 
                         tubbs_bert$score)

colnames(tubbs_tweets)[8] <- "bert_sent"
colnames(tubbs_tweets)[9] <- "bert_score"

# 2334 Unique Users
tubbs_users <- tubbs_users %>% 
  group_by(author_id) %>% 
  summarize_all(max)

# Join into one table
tubbs <- inner_join(tubbs_tweets, tubbs_users, by = "author_id") %>%
  mutate(fire = "Tubbs Fire")
```

```{r}
# Create Main Data Frame
main <- rbind(bay, camp, dixie, tubbs)

# Keep a copy of original tweet
main$orig_text <- main$text
```

```{r}
# Pre-Processing
## Remove links and websites
main$text <- gsub("https.*","",main$text)
main$text <- gsub("http.*","",main$text)
main$text <- gsub("www.*","",main$text)
## Remove Twitter Handles
main$text <- gsub("@([a-zA-Z0-9]|[_])*","",main$text)
```

```{r}
# Run Sentiment from syuzhet package for pre-processed text
syuzhet_df <- get_sentiment(main$text) %>% as.data.frame()
```

```{r}
# Add syuzhet sentiment score to main dataframe
main <- cbind(main, syuzhet_df)
colnames(main)[19] <- "syuzhet_score"
```

```{r}
# Syuzhet Range: -7.10 to 6.35
# > 0 is Positive
# < 0 is Negative
# = 0 is Neutral
# Add syuzhet sentiment score classification

main <- main %>% 
  mutate(syuzhet_sent = case_when(syuzhet_score > 0 ~ "positive", 
                               syuzhet_score < 0 ~ "negative",
                               syuzhet_score == 0 ~ "neutral"))
```

```{r}
# Save DataFrame as CSV
write.csv(main, "MainSentiment.csv", row.names = FALSE)
```

**Syuzhet**

```{r}
main %>%
  filter(followers_count > 0 & 
           following_count > 0 & 
           like_count > 0) %>%
  filter(!stringr::str_detect(description,'news')) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(Positive = length(syuzhet_sent[syuzhet_sent == "positive"]) / n(),
            Negative = length(syuzhet_sent[syuzhet_sent == "negative"]) / n(),
            Neutral = length(syuzhet_sent[syuzhet_sent == "neutral"]) / n(),
            Total = n()) %>%
  gather(sentiment, percentage, Positive:Neutral) %>%
  mutate(
    se = sqrt(percentage * (1 - percentage) / Total),
    lower = percentage - 2 * se,
    upper = percentage + 2 * se) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.3) + 
  geom_errorbar(aes(ymin = lower, ymax = upper)
                , position = "dodge"
                , color = "black") +
  scale_fill_manual(values = c("red2", "blue", "darkgreen")) +
  labs(title = "Syuzhet Sentiment Analysis",
       x = "Fire",
       y = "Proportion of Tweets",
       fill = "Sentiment") +
  theme_bw()
```

**BERT**

```{r}
 main %>%
  filter(followers_count > 0 & 
           following_count > 0 & 
           like_count > 0) %>%
  filter(!stringr::str_detect(description,'news')) %>%
  mutate(fire = fct_relevel(fire, c("Tubbs Fire", 
                      "Camp Fire", 
                      "Bay Area Fire", 
                      "Dixie Fire"))) %>%
  group_by(fire) %>%
  summarize(Positive = length(bert_sent[bert_sent == "positive"]) / n(),
            Negative = length(bert_sent[bert_sent == "negative"]) / n(),
            Neutral = length(bert_sent[bert_sent == "neutral"]) / n(),
            Total = n()) %>%
  gather(sentiment, percentage, Positive:Neutral) %>%
  mutate(
    se = sqrt(percentage * (1 - percentage) / Total),
    lower = percentage - 2 * se,
    upper = percentage + 2 * se) %>%
  ggplot(aes(x = fire, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.3) + 
  geom_errorbar(aes(ymin = lower, ymax = upper)
                , position = "dodge"
                , color = "black") +
  scale_fill_manual(values = c("red2", "blue", "darkgreen")) +
  labs(title = "BERT Sentiment Analysis",
       x = "Fire",
       y = "Proportion of Tweets",
       fill = "Sentiment") +
  theme_bw()
```